---
title: "State and Local Government Finance Research Lab"
subtitle: "Advancing empirical research in public finance"
author: "Led by Professor Gang Chen"
date: "`r Sys.Date()`"
format: 
  html:
    theme: 
      - cosmo
      - custom.scss
    toc: true
    toc-location: left
    toc-title: "Navigation"
    css: styles.css
    include-in-header: |
      <link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/6.0.0/css/all.min.css">
    page-layout: full
    grid:
      sidebar-width: 250px
      body-width: 800px
      margin-width: 150px
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(knitr)
library(DT)
library(plotly)
library(dplyr)
```

::: {.hero-section}
## About SLGF-Lab

**State and Local Government Finance Research Lab (SLGF-Lab)** is a research branch of the [State and Local Government Finance (SLGF) Project](https://www.albany.edu/rockefeller/project-state-local-government-finance) at the Center for Policy Research at Rockefeller College, University at Albany SUNY. 

SLGF-Lab is led by **Professor Gang Chen**, Co-Director of the SLGF Project. The lab focuses on empirical research on state and local government budgeting and financial management, using advanced econometric methods, simulations, and large-scale datasets.

::: {.callout-note icon=false}
## Research Status
The research presented on this website represents ongoing work conducted in the lab. Full reports will be published once research is completed.
:::
:::

## Current Research Projects {#projects}

### Municipal Tax Override Studies {#municipal-tax}

::: {.project-card}
**Property Tax Overrides and Municipal Credit Ratings: Evidence from Massachusetts**

::: {.project-meta}
<i class="fas fa-calendar-alt"></i> **Timeline:** 2003-2021 data  
<i class="fas fa-map-marker-alt"></i> **Location:** Massachusetts municipalities  
<i class="fas fa-chart-line"></i> **Status:** Analysis phase
:::

A municipal tax override is a voter-approved measure that allows a city or town to temporarily raise property taxes beyond statutory limits. While overrides enable governments to generate additional revenue in response to rising public service costs, they may also undermine the purpose of statutory limits to promote fiscal sustainability. 

**Research Methodology:**
- Comprehensive dataset from Massachusetts municipalities (2003-2021)
- Econometric analysis of override impacts
- Credit rating assessment models
- Voter behavior analysis

**Key Findings:**
- ✅ Successful overrides associated with improved credit ratings
- ⚠️ Frequent overrides can damage credit ratings
- ❌ Failed override attempts negatively impact future voter support

[View Research Details](#municipal-details){.btn .btn-primary}
:::

### Rural School Finance Studies {#rural-schools}

::: {.project-card}
**Examining the Impact of Legacy Costs on New York State Rural Schools**

::: {.project-meta}
<i class="fas fa-graduation-cap"></i> **Focus:** Rural school districts  
<i class="fas fa-map-marker-alt"></i> **Location:** Upstate New York  
<i class="fas fa-dollar-sign"></i> **Key Issue:** Legacy cost burden
:::

This research focuses on the growing burden of "legacy costs" (pensions, retiree healthcare, and capital project maintenance) in rural school districts facing declining enrollments. These costs consume an increasing share of budgets, crowding out instructional spending and straining schools' financial capacity.

**Research Components:**
- Quantification of legacy cost trends
- Rural vs. non-rural district comparisons  
- Budget impact assessments
- Operational capacity analysis

**Data Sources:**
- New York State comprehensive school financial data
- Enrollment trend analysis
- Multi-year budget comparisons

[View Research Details](#rural-details){.btn .btn-primary}
:::

::: {.project-card}
**Examining the Causes and Status of Fiscal Stress in Rural Schools**

::: {.project-meta}
<i class="fas fa-exclamation-triangle"></i> **Focus:** Fiscal stress indicators  
<i class="fas fa-balance-scale"></i> **Analysis:** Comprehensive financial metrics  
<i class="fas fa-users"></i> **Impact:** Policy recommendations
:::

Rural school districts in New York face shrinking enrollments, limited tax bases, and rising fixed costs, putting them at heightened risk of fiscal stress. This study analyzes the status, causes, and consequences of fiscal stress using comprehensive financial indicators.

**Financial Indicators Analyzed:**
- Fund balance ratios
- Liquidity measures  
- Debt service ratios
- Revenue diversification metrics
- Expenditure per pupil trends

**Policy Implications:**
The findings inform policymakers and educators about fiscal stress drivers, contributing to policy solutions aimed at reducing inequalities and improving educational equity across New York State.

[View Research Details](#fiscal-stress-details){.btn .btn-primary}
:::

## Research Methods & Data {#methods}

::: {.methodology-section}
### Our Approach

SLGF-Lab employs rigorous empirical methods to ensure robust and reliable research outcomes:

::: {.method-grid}
::: {.method-item}
**<i class="fas fa-calculator"></i> Econometric Analysis**
- Panel data models
- Causal inference techniques  
- Regression discontinuity designs
- Difference-in-differences estimation
:::

::: {.method-item}
**<i class="fas fa-database"></i> Large-Scale Datasets**
- Multi-year municipal financial data
- School district comprehensive databases
- Credit rating agency records
- Voter behavior datasets
:::

::: {.method-item}
**<i class="fas fa-cogs"></i> Simulation Models**
- Fiscal impact projections
- Policy scenario analysis
- Budget stress testing
- Revenue forecasting models
:::
:::
:::

```{r}
#| label: research-timeline
#| fig-height: 4

# Create a simple research timeline visualization
timeline_data <- data.frame(
  Project = c("MA Tax Override Study", "NY Rural Schools - Legacy Costs", "NY Rural Schools - Fiscal Stress"),
  Start_Year = c(2022, 2023, 2023),
  Expected_Completion = c(2025, 2025, 2024),
  Status = c("Analysis", "Data Collection", "Analysis")
)

# Create timeline plot
p <- timeline_data %>%
  plotly::plot_ly(
    x = ~Start_Year, 
    y = ~Project,
    xend = ~Expected_Completion,
    yend = ~Project,
    mode = 'lines+markers',
    type = 'scatter',
    line = list(width = 6),
    marker = list(size = 10)
  ) %>%
  plotly::add_segments(
    x = ~Start_Year, 
    xend = ~Expected_Completion,
    y = ~Project, 
    yend = ~Project,
    line = list(width = 6, color = 'steelblue')
  ) %>%
  plotly::layout(
    title = "Research Project Timeline",
    xaxis = list(title = "Year"),
    yaxis = list(title = ""),
    showlegend = FALSE
  )

p
```

## Lab Team & Collaboration {#team}

::: {.team-section}
### Leadership

**Professor Gang Chen**  
*Lab Director & Co-Director of SLGF Project*  
- Specialization: State and local public finance
- Email: [gchen@albany.edu](mailto:gchen@albany.edu)  
- Office: Rockefeller College, University at Albany SUNY

### Research Opportunities

We welcome collaboration with:
- Graduate students interested in public finance research
- Fellow researchers at other institutions  
- Policy practitioners and government officials
- Data partners and funding agencies

For collaboration inquiries, please contact the lab director.
:::

## Publications & Resources {#publications}

::: {.publications-section}
### Working Papers
*Coming soon - manuscripts currently in preparation*

### Data Resources
- Municipal financial indicators dashboard *(in development)*
- School district fiscal health monitoring system *(in development)*
- Interactive policy simulation tools *(planned)*

### Policy Briefs
Research findings will be disseminated through accessible policy briefs for practitioners and policymakers.
:::

## Contact Information {#contact}

::: {.contact-section}
**State and Local Government Finance Research Lab**  
Center for Policy Research  
Rockefeller College of Public Affairs & Policy  
University at Albany, SUNY  

**Mailing Address:**  
135 Western Ave  
Albany, NY 12222  

**Email:** [slgf-lab@albany.edu](mailto:slgf-lab@albany.edu)  
**Web:** [www.albany.edu/rockefeller/slgf-lab](https://www.albany.edu/rockefeller/slgf-lab)  

---

*This website is maintained by SLGF-Lab. Last updated: `r format(Sys.Date(), "%B %d, %Y")`*
:::

```{css}
/* Custom CSS for enhanced styling */

.hero-section {
  background: linear-gradient(135deg, #f8fafc 0%, #e2e8f0 100%);
  padding: 2rem;
  border-radius: 10px;
  margin-bottom: 2rem;
  border-left: 5px solid #3b82f6;
}

.project-card {
  background: white;
  padding: 2rem;
  border-radius: 8px;
  box-shadow: 0 4px 6px rgba(0, 0, 0, 0.1);
  margin-bottom: 2rem;
  border-left: 4px solid #f59e0b;
  transition: transform 0.2s ease-in-out;
}

.project-card:hover {
  transform: translateY(-2px);
  box-shadow: 0 8px 15px rgba(0, 0, 0, 0.15);
}

.project-meta {
  background-color: #f8fafc;
  padding: 1rem;
  border-radius: 6px;
  margin: 1rem 0;
  font-size: 0.9rem;
}

.project-meta i {
  color: #3b82f6;
  margin-right: 0.5rem;
}

.btn {
  display: inline-block;
  padding: 0.5rem 1rem;
  background-color: #3b82f6;
  color: white;
  text-decoration: none;
  border-radius: 5px;
  transition: background-color 0.3s ease;
  margin-top: 1rem;
}

.btn:hover {
  background-color: #2563eb;
  color: white;
  text-decoration: none;
}

.methodology-section {
  background-color: #f8fafc;
  padding: 2rem;
  border-radius: 8px;
  margin: 2rem 0;
}

.method-grid {
  display: grid;
  grid-template-columns: repeat(auto-fit, minmax(250px, 1fr));
  gap: 1.5rem;
  margin-top: 1.5rem;
}

.method-item {
  background: white;
  padding: 1.5rem;
  border-radius: 6px;
  box-shadow: 0 2px 4px rgba(0, 0, 0, 0.1);
}

.method-item i {
  color: #f59e0b;
  font-size: 1.5rem;
  margin-bottom: 0.5rem;
}

.team-section, .publications-section, .contact-section {
  background-color: #f9fafb;
  padding: 2rem;
  border-radius: 8px;
  margin: 2rem 0;
}

.callout-note {
  background-color: #dbeafe;
  border-left: 4px solid #3b82f6;
  padding: 1rem 1.5rem;
  border-radius: 0 6px 6px 0;
}

/* Responsive design */
@media (max-width: 768px) {
  .method-grid {
    grid-template-columns: 1fr;
  }
  
  .project-card {
    padding: 1.5rem;
  }
  
  .hero-section {
    padding: 1.5rem;
  }
}
```